---
title: "p8105_hw1_nn2477"
author: "Nhu Nguyen"
date: "2023-09-15"
output: github_document
---

**Problem 1** 
```{r}
library("moderndive")
data("early_january_weather")

#finding the variables in dataset
names(early_january_weather)
early_january_weather

#display total number of rows in dataset
nrow(early_january_weather)
```
There are 358 rows in the early_january_weather dataset 

```{r}
#display total number of columns in dataset
ncol(early_january_weather)
```
There are 15 columns in the early_january_weather dataset 

```{r}
#finding mean for variable temp
mean(early_january_weather$temp)
```
The mean temperature is 39.58. 

**Creating scatter plot and saving graph**
```{r}
#load ggplot2 library
library(ggplot2)

#making scatterplot using ggplot function 
scatter_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + 
  geom_point() +
  labs(
    x = "Time (Hours)",
    y = "Temperature",
    title = "Scatterplot of Temperature vs Time with Humidity Color Code"
  ) + 
  scale_color_gradient(low = "blue", high = "red") + 
  theme_minimal()

#print scatterplot 
print(scatter_plot)

ggsave("scatterplot.png", plot = scatter_plot)
```

**Scatter plot description:**
there seems to be an overall positive correlation, meaning that as time (hrs) increases, temperature also increases as well. however, no clear trend is evident, indicating a weak correlation between time and temperature


