---
title: "p8105_hw1_nn2477"
author: "Nhu Nguyen"
date: "2023-09-15"
output: github_document
---

**Problem 1** 
```{r}
library(tidyverse)
library("moderndive")
data("early_january_weather")

#finding the variables in data set
names(early_january_weather)
early_january_weather

#display total number of rows in dataset
nrow(early_january_weather)

#display total number of columns in dataset
ncol(early_january_weather)

#finding mean for variable temp
mean(early_january_weather$temp)

#finding the sd for variable temp 
sd(early_january_weather$temp)
```


**Data set description:** This data set contains weather factors for the first half of January 2013. It has 358 observations and 15 variables, including dew point, humidity, wind characteristics, precipitation, and pressure. The mean temperature for the first half of January 2013 was 39.58 degrees Fahrenheit with a standard deviation of 7.06. 


**Creating scatter plot and saving graph**
```{r}
#load ggplot2 library
library(ggplot2)

#making scatterplot using ggplot function 
scatter_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + 
  geom_point() +
  labs(
    x = "Time (Hours)",
    y = "Temperature",
    title = "Scatterplot of Temperature vs Time with Humidity Color Code"
  ) + 
  scale_color_gradient(low = "blue", high = "red") + 
  theme_minimal()

#print scatterplot 
print(scatter_plot)

```

```{r}
#save scatter plot as png
ggsave("scatterplot.png", plot = scatter_plot)
```


**Scatter plot description:**
there seems to be an overall positive correlation, meaning that as time (hrs) increases, temperature also increases as well. however, no clear trend is evident, indicating a weak correlation between time and temperature

**Problem 2**
```{r}
#creating new data frame
df = tibble(
  random_sample = rnorm(10), 
  log_vec = rnorm(10) > 0, 
  char_vec = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  fact_vec = factor(sample(1:3, 10, replace = TRUE), levels = 1:3, labels = c("Level_1", "Level_2", "Level_3"))
)

#checking to see if data frame is created correctly 
print(df)

```




