---
title: "p8105_hw1_nn2477"
author: "Nhu Nguyen"
date: "2023-09-15"
output: github_document
---

**Problem 1** 
```{r}
library(tidyverse)
library("moderndive")
data("early_january_weather")

#finding the variables in data set
names(early_january_weather)
early_january_weather

#display total number of rows in dataset
nrow(early_january_weather)

#display total number of columns in dataset
ncol(early_january_weather)

#finding mean for variable temp
mean(early_january_weather$temp)

#finding the sd for variable temp 
sd(early_january_weather$temp)
```


**Data set description:** This data set contains weather factors for the first half of January 2013. It has 358 observations and 15 variables, including dew point, humidity, wind characteristics, precipitation, and pressure. The mean temperature for the first half of January 2013 was 39.58 degrees Fahrenheit with a standard deviation of 7.06. 


**Creating scatter plot and saving graph**
```{r}
#load ggplot2 library
library(ggplot2)

#making scatterplot using ggplot function 
scatter_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + 
  geom_point() +
  labs(
    x = "Time (Hours)",
    y = "Temperature",
    title = "Scatterplot of Temperature vs Time with Humidity Color Code"
  ) + 
  scale_color_gradient(low = "blue", high = "red") + 
  theme_minimal()

#print scatterplot 
print(scatter_plot)

```

```{r}
#save scatter plot as png
ggsave("scatterplot.png", plot = scatter_plot)
```


**Scatter plot description:**
there seems to be an overall positive correlation, meaning that as time (hrs) increases, temperature also increases as well. however, no clear trend is evident, indicating a weak correlation between time and temperature

**Problem 2**
```{r}
#creating new data frame
df = tibble(
  random_sample = rnorm(10), 
  logical_vector = rnorm(10) > 0, 
  character_vector = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  factor_vector = factor(sample(1:3, 10, replace = TRUE), levels = 1:3, labels = c("Low", "Medium", "High"))
)

#checking to see if data frame is created correctly 
print(df)

#find mean for each variable using pull()
mean_random_sample = mean(pull(df, random_sample))
print(mean_random_sample)

mean_log_vec = mean(pull(df, logical_vector))
print(mean_logical_vector)

mean_char_vec = mean(pull(df, character_vector))
print(mean_character_vector)

mean_fact_vec = mean(pull(df, factor_vector))
print(mean_factor_vector)
```

I was able to take the mean for the variables "random_sample" and "logical_vector" because it contains numeric and/or logical values. I was unable to take the mean for variables "character_vector" and "factor_vector" because it is not a numeric and/or logical outputs. 


```{r}
#apply as.numeric() to logical, character, and factor variables 
numeric_logical <- as.numeric(df$logical_vector)
```


Using the as.numeric() function on the logical vector converts 'true' values to 1 and 'false' values to 0. This conversion makes sense because 'true' and 'false' are usually treated as 1 and 0 in many data sets. This explains why when we take the mean of the "logical_vector" variable, our output is 0.5 as that is the average. 


```{r}
numeric_character <- as.numeric(df$character_vector)
```

Using the as.numeric() function on the character vector in R, it tries to convert the character values to numeric values. However, since our values are non-numeric characters, R converts them to missing values (NAs). We are unable to find the mean of the "character_vector" variable because of the presence of NAs (it would just give us an NA), thus it explains why we get an NA as the output when we try to find the mean for this variable.


```{r}
numeric_factor <- as.numeric(df$factor_vector)
```

Using the as.numeric() function on the factor vector in R, it attempts to convert the factor levels based on their internal encoding and these numeric values may not align with the intended factor order. This helps explain why I get an NA when attempting to find the mean since factors are categorical variables and their levels are represented as integers internally.